Classification Prediction with using Weight lifting data
========================================================


## Background


The goal of this analysis is to predict the manner in which of the 6 participants 
did the exercise( i.e., the "classe" variable) using data from accelerometers on the belt, forearm, arm, and dumbell of 6 participants. They were asked to perform barbell lifts correctly and incorrectly in 5 different ways.


## Cleaning and Preparing Data
Download data from the course website and load the data to R

```{r}
URL<-"http://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
download.file(URL,destfile="training.csv")
train = read.csv("training.csv")
```



After a quick look of the data, we find there are a lot missing values(NA, NaN and inf) in the dataset. 
Cleaning data
```{r echo =TRUE}
library(caret)
missing_index = nearZeroVar(train)
train = train[-missing_index]
features = which(lapply(train, class) %in% c("numeric"))
```



Divide the dataset into to parts: 70% for training and 30% for testing.


```{r,cache=TRUE}
train_index<-createDataPartition(y = train$classe, p=0.3,list=FALSE) 
training<-train[train_index,]
testing<-train[-train_index,]
```



preprocess the data
```{r echo=TRUE}
preProc <- preProcess(training[,features], method = "knnImpute")
trainingProc <- predict(preProc, training[,features])
trainingProc<- cbind( "classe" = training$classe, trainingProc)
testingProc <- predict(preProc, testing[,features])
testingProc <- cbind(  "classe"=testing$classe, testingProc)
```





## machine learning algorithms  
#### random forest, building the model


```{r echo =TRUE}
library(randomForest)
rf.fit <- randomForest(classe ~., data = trainingProc, ntree = 800, mtry = 20)

```

predict the model with the training and testing data (from the original training data)

```{r echo=TRUE}
rf.predict <- predict(rf.fit, testingProc)
rf.table <- confusionMatrix(testingProc$classe, rf.predict)
rf.table
```

#### lda
```{r echo=TRUE, results='hide'}
library(MASS)
lda.fit <- train(classe ~ ., data = trainingProc, method = "lda")
```

predict
```{r echo=TRUE}
lda.predict <- predict(lda.fit, testingProc)
lda.table <- confusionMatrix(testingProc$classe, lda.predict)
lda.table
```




# Estimate Out Of Sample Error
After comparing the accuracy reported from the above two models, random forest model is selected as the final model. 
Apply model to the test dataset

```{r eval=FALSE, echo=TRUE, results='hide'}
test = read.csv("test.csv")
test.predict<- predict(rf.fit, test)
test.table<- confusionMatrix(test$classe, test.predict)
print(test.table$overall[1])
```
Accuracy 
0.9702


So the estimated out of sample error rate is **2.98%**
